---
title: "data_prep"
author: "CEB"
date: "1/31/2022"
output: html_document
---

####---------Load and clean seabird data------------

##load packages
```{r}
library(tidyverse)
library(vegan)
```


##Load files - Chagos 2015 only
```{r}
seabird_meta<-read.csv("seabird_meta_chagos_only_2021_09_30.csv", stringsAsFactors = TRUE)
seabird_dat<-read.csv("seabird_data_combined_chagos_only.csv", stringsAsFactors = TRUE)

str(seabird_meta)
str(seabird_dat)

```


##Combine seabird count data with seabird meta data -  ((breeding pairs x 2 x weight x (months present/12))/ area for annual biomass 
```{r}
##first get proportion of year present at each island
seabird_meta <-
  seabird_meta %>%
  mutate(prop_year = months_chagos/12)
seabird_meta

##change scientific name to match rows in data sheet:
seabird_meta <-
  seabird_meta %>%
  mutate(Scientific.name = str_trim(Scientific.name, side = "both")) %>%  ###remove trailing spaces....
  mutate(Scientific.name = as.factor(str_replace_all(Scientific.name, " ", ".")))


###merge data
seabird_dat_ch_2015_long<-
seabird_dat %>% 
    pivot_longer(c(Ardenna.pacifica:Gygis.alba), names_to="Scientific.name", values_to="Num_Breeding_Pairs")  %>% 
    mutate(Scientific.name = as.factor(Scientific.name))
seabird_dat_ch_2015_long


seabird_dat_ch_2015_meta <-
  inner_join(seabird_dat_ch_2015_long, seabird_meta, by = "Scientific.name")
###okay, now one row per species per island :)
seabird_dat_ch_2015_meta

```


##now do calculations - biomass, per hectare
```{r}

seabird_dat_ch_2015_bio<-
  seabird_dat_ch_2015_meta %>% 
  mutate(Total_Biomass_kg = Num_Breeding_Pairs*2*weight_g/1000) %>% #divide by 1000 because g to kg
  mutate(Biomass_Prop_Year_kg = Total_Biomass_kg*prop_year)%>% 
  mutate(Biomass_Prop_Year_kg_ha= (Biomass_Prop_Year_kg)/island_size_ha)%>% 
  mutate(Breeding_Pairs_Prop_Year = Num_Breeding_Pairs*prop_year)%>% 
  mutate(Breeding_Pairs_Prop_Year_ha = Breeding_Pairs_Prop_Year/island_size_ha)%>% 
  mutate(Breeding_Pairs_ha = Num_Breeding_Pairs/island_size_ha)%>% 
  mutate(Biomass_kg_ha = Total_Biomass_kg/island_size_ha)%>% 
  select(Region, Atoll, Island, rat_status, island_size_ha,
         Family, Scientific.name, Num_Breeding_Pairs, 
         Total_Biomass_kg, Biomass_Prop_Year_kg, Biomass_Prop_Year_kg_ha, 
         Breeding_Pairs_Prop_Year,  Breeding_Pairs_Prop_Year_ha, 
         Breeding_Pairs_ha, Biomass_kg_ha) 

seabird_dat_ch_2015_bio

seabird_dat_ch_2015_bio<-
  seabird_dat_ch_2015_bio%>% 
      mutate(Biomass_Prop_Year_kg = round(Biomass_Prop_Year_kg, 4)) %>%
    mutate(Biomass_Prop_Year_kg_ha = round(Biomass_Prop_Year_kg_ha, 4)) %>%
  mutate(Breeding_Pairs_Prop_Year = round(Breeding_Pairs_Prop_Year, 4)) %>%
  mutate(Breeding_Pairs_Prop_Year_ha = round(Breeding_Pairs_Prop_Year_ha, 4))  %>%
  mutate(Breeding_Pairs_ha = round(Breeding_Pairs_ha, 4)) %>%
mutate(Biomass_kg_ha = round(Biomass_kg_ha, 4)) 

seabird_dat_ch_2015_bio
  
```


###sum by island
```{r}

###by island only
seabird_dat_ch_2015_bio_isl<-
  seabird_dat_ch_2015_bio %>% 
  group_by(Island, rat_status) %>% 
   summarize(sum_bio_total_kg = sum(Total_Biomass_kg),
             sum_bio_prop_year_kg = sum(Biomass_Prop_Year_kg),
            sum_bio_prop_year_kg_ha = sum(Biomass_Prop_Year_kg_ha),
            sum_breeding_pairs = sum(Num_Breeding_Pairs),
            sum_breeding_pairs_prop_year = sum(Breeding_Pairs_Prop_Year),
            sum_breeding_pairs_prop_year_ha = sum(Breeding_Pairs_Prop_Year_ha),
            sum_breeding_pairs_ha = sum(Breeding_Pairs_ha),
            sum_biomass_kg_ha = sum(Biomass_kg_ha))

seabird_dat_ch_2015_bio_isl


###so dataframes for species, island respectively:
#seabird_dat_ch_2015_bio
#seabird_dat_ch_2015_bio_isl


##add row columns with island totals to species data:
seabird_chagos_2015_dat_meta<-left_join(seabird_dat_ch_2015_bio, seabird_dat_ch_2015_bio_isl, by = c("Island", "rat_status"))
seabird_chagos_2015_dat_meta

```


##make species data wide
```{r}

#do biomass and breeding pairs together:

ch_15_bio_wide<-
  seabird_chagos_2015_dat_meta  %>% 
  pivot_wider(id_cols =c("Atoll", "Island", "rat_status", "island_size_ha",
                         "sum_bio_prop_year_kg_ha"), 
              names_from = "Scientific.name", values_from = c("Biomass_Prop_Year_kg_ha", "Breeding_Pairs_ha"))


```


###load and combine isotope data
```{r}
##combine with isotope data
##load nutrient data
iso_dat_ch15<-read.csv("stable_istotope_2015chagos_2020_05_06.csv", stringsAsFactors = TRUE) 

#check island levels
levels(iso_dat_ch15$Island)
levels(ch_15_bio_wide$Island)

#change island levels to make match, so then can combine: 
iso_dat_ch_rl<-
  iso_dat_ch15%>% 
  mutate(Island = recode_factor(Island, "Eagle"  ="Eagle Island", "Grande_Ile_Coquillage"= "Grande Ile Coquillage",  
                                "Grande_Ile_Mapou" ="Grande Ile Mapou" ,    "PB_Ile_Anglaise"   = "Ile Anglaise_PB" ,     
                                "Sal_Ile_Anglaise"   =  "Ile Anglaise_SAL"  ,    "Ile_de_la_Passe" = "Ile de la Passe",     
                                "Ile_Fouquet"   =  "Ile Fouquet" ,"Ile_Longue"  ="Ile Longue"   ,        "Ile_Poule"    =  "Ile Poule" ,   
                                "Middle_Brother"  =   "Middle Brother"  ,    "Nelson_Island"   =  "Nelson's Island" ,   
                                "South_Brother"   =   "South Brother" ))


str(ch_15_bio_wide)
str(iso_dat_ch_rl)

##combine:
iso_bio_ch<-inner_join(iso_dat_ch_rl, ch_15_bio_wide,  by=c("Island")) 
iso_bio_ch
```



##get diversity metrics
```{r}

str(ch_15_bio_wide)

#create data for habitat types:
sb_types<-
  ch_15_bio_wide%>% 
  select(Biomass_Prop_Year_kg_ha_Ardenna.pacifica:Biomass_Prop_Year_kg_ha_Gygis.alba)

ch_2015_div<-
  ch_15_bio_wide%>% 
  mutate(richness = specnumber(sb_types)) %>% 
  mutate(log_bio_kg_ha = log(sum_bio_prop_year_kg_ha),
         log_rich = log(richness),
         rich_cs = scale(richness, center = TRUE, scale = TRUE),
         logbio_cs = scale(log_bio_kg_ha, center = TRUE, scale = TRUE),
         logrich_cs = scale(log_rich, center = TRUE, scale = TRUE),
         logbio_c = scale(log_bio_kg_ha, center = TRUE, scale = FALSE),
         rich_c = scale(richness, center = TRUE, scale = FALSE))


#combine data
ch_iso_div_15<- inner_join(iso_dat_ch_rl, ch_2015_div, by = "Island")
ch_iso_div_15

```


##save as Rdata files
```{r}

##seabird/site data:
seabird_dat_ch_2015_meta

#with calculations:
seabird_dat_ch_2015_bio

#by island (but no diversity stats):
seabird_dat_ch_2015_bio_isl

#by species with island totals:
seabird_chagos_2015_dat_meta

#by island, wide with diversity data:
ch_2015_div

##seabird/site data wide:
ch_15_bio_wide #biomass (prop year kg/ha) and breeding pairs (#/ha)

#isotope data: 
iso_dat_ch_rl

##isotope plus seabird data:
iso_bio_ch

##isotope plus diversity data
ch_iso_div_15

save(seabird_dat_ch_2015_meta, seabird_dat_ch_2015_bio, seabird_dat_ch_2015_bio_isl, 
     seabird_chagos_2015_dat_meta, ch_2015_div,
     ch_15_bio_wide,
     iso_dat_ch_rl,iso_bio_ch, ch_iso_div_15,
     file = "seabird_comm_data_2021_09_30.Rdata")
```


###-------create other datasheet where we eliminate rare species---------

#remove rare species
```{r}
seabird_dat_ch_2015_meta_REDUCED3<-
  seabird_dat_ch_2015_meta %>%
  filter(Scientific.name != "Phaethon.lepturus" )  %>%
  filter(Scientific.name != "Sterna.dougallii") %>%
  droplevels()

levels(seabird_dat_ch_2015_meta_REDUCED3$Scientific.name)

```


##now do calculations with reduced data, plus add in isotope data
```{r}

seabird_dat_ch_2015_bio_STDO_and_PHLE_removed<-
  seabird_dat_ch_2015_meta_REDUCED3 %>% 
  mutate(Total_Biomass_kg = Num_Breeding_Pairs*2*weight_g/1000) %>% #divide by 1000 because g to kg
  mutate(Biomass_Prop_Year_kg = Total_Biomass_kg*prop_year)%>% 
  mutate(Biomass_Prop_Year_kg_ha= (Biomass_Prop_Year_kg)/island_size_ha)%>% 
  mutate(Breeding_Pairs_Prop_Year = Num_Breeding_Pairs*prop_year)%>% 
  mutate(Breeding_Pairs_Prop_Year_ha = Breeding_Pairs_Prop_Year/island_size_ha)%>% 
  mutate(Breeding_Pairs_ha = Num_Breeding_Pairs/island_size_ha)%>% 
  mutate(Biomass_kg_ha = Total_Biomass_kg/island_size_ha)%>% 
  select(Region, Atoll, Island, rat_status,  island_size_ha, 
         Family, Scientific.name, Num_Breeding_Pairs, 
         Total_Biomass_kg, Biomass_Prop_Year_kg, Biomass_Prop_Year_kg_ha, 
         Breeding_Pairs_Prop_Year,  Breeding_Pairs_Prop_Year_ha, 
         Breeding_Pairs_ha, Biomass_kg_ha) 

#round decimal places to 4
seabird_dat_ch_2015_bio_STDO_and_PHLE_removed<-
  seabird_dat_ch_2015_bio_STDO_and_PHLE_removed%>% 
      mutate(Biomass_Prop_Year_kg = round(Biomass_Prop_Year_kg, 4)) %>%
    mutate(Biomass_Prop_Year_kg_ha = round(Biomass_Prop_Year_kg_ha, 4)) %>%
  mutate(Breeding_Pairs_Prop_Year = round(Breeding_Pairs_Prop_Year, 4)) %>%
  mutate(Breeding_Pairs_Prop_Year_ha = round(Breeding_Pairs_Prop_Year_ha, 4))  %>%
  mutate(Breeding_Pairs_ha = round(Breeding_Pairs_ha, 4)) %>%
mutate(Biomass_kg_ha = round(Biomass_kg_ha, 4)) 

seabird_dat_ch_2015_bio_STDO_and_PHLE_removed
  
```

###sum by island
```{r}

###by island only
seabird_dat_ch_2015_bio_STDO_and_PHLE_removed_isl<-
  seabird_dat_ch_2015_bio_STDO_and_PHLE_removed %>% 
  group_by(Island, rat_status) %>% 
   summarize(sum_bio_total_kg = sum(Total_Biomass_kg),
             sum_bio_prop_year_kg = sum(Biomass_Prop_Year_kg),
            sum_bio_prop_year_kg_ha = sum(Biomass_Prop_Year_kg_ha),
            sum_breeding_pairs = sum(Num_Breeding_Pairs),
            sum_breeding_pairs_prop_year = sum(Breeding_Pairs_Prop_Year),
            sum_breeding_pairs_prop_year_ha = sum(Breeding_Pairs_Prop_Year_ha),
            sum_breeding_pairs_ha = sum(Breeding_Pairs_ha),
            sum_biomass_kg_ha = sum(Biomass_kg_ha))

##add row columns with island totals to species data:
seabird_dat_ch_2015_bio_meta_R3<-left_join(seabird_dat_ch_2015_bio_STDO_and_PHLE_removed, seabird_dat_ch_2015_bio_STDO_and_PHLE_removed_isl, by = c("Island", "rat_status"))
seabird_dat_ch_2015_bio_meta_R3

```


##make species data wide 
```{r}
#do biomass and breeding pairs together:
ch_15_bio_wide_R3<-
  seabird_dat_ch_2015_bio_meta_R3  %>% 
  pivot_wider(id_cols =c("Atoll", "Island", "rat_status",  "island_size_ha",
                         "sum_bio_prop_year_kg_ha"), 
              names_from = "Scientific.name", values_from = c("Biomass_Prop_Year_kg_ha", "Breeding_Pairs_ha"))

```


##and now combine reduced data with isotope data:
```{r}
iso_bio_ch

##combine:
iso_bio_ch_R3<-inner_join(iso_dat_ch_rl, ch_15_bio_wide_R3,  by=c("Island")) 
iso_bio_ch_R3

```


##finally, get diversity metrics and combine
```{r}

sb_types_R3<-
  ch_15_bio_wide_R3%>% 
  select(Biomass_Prop_Year_kg_ha_Ardenna.pacifica:Biomass_Prop_Year_kg_ha_Gygis.alba)

ch_2015_div_R3<-
  ch_15_bio_wide_R3%>% 
  mutate(richness = specnumber(sb_types_R3)) %>% 
  mutate(log_bio_kg_ha = log(sum_bio_prop_year_kg_ha),
         log_rich = log(richness),
         rich_cs = scale(richness, center = TRUE, scale = TRUE),
         logbio_cs = scale(log_bio_kg_ha, center = TRUE, scale = TRUE),
         logrich_cs = scale(log_rich, center = TRUE, scale = TRUE),
         logbio_c = scale(log_bio_kg_ha, center = TRUE, scale = FALSE),
         rich_c = scale(richness, center = TRUE, scale = FALSE))


#combine data
ch_iso_div_15_R3<- inner_join(iso_dat_ch_rl, ch_2015_div_R3, by = "Island")
ch_iso_div_15_R3

```

###save as .rdata
```{r}

save(ch_2015_div_R3, ch_iso_div_15_R3,seabird_dat_ch_2015_bio_meta_R3,
     file = "seabird_comm_data_REDUCED_2021_09_30.Rdata")
```



######------------combine seabird and vegetation data--------------------------


##load data
```{r}
veg_dat<-read.csv("island_vegetation_dat.csv", stringsAsFactors = TRUE)

levels(veg_dat$Island)
levels(ch_2015_div$Island)

levels(veg_dat$Atoll)
levels(ch_2015_div$Atoll)

ch_2015_bio_div_veg<-inner_join(veg_dat, ch_2015_div,  by=c("Atoll","Island")) 
ch_2015_bio_div_veg
str(ch_2015_bio_div_veg)

#reduce to what we need:
ch_2015_sb_veg<-
  ch_2015_bio_div_veg%>% 
  select(c(Island, Atoll, rat_status, island_size_ha, non.native_forest_percent,  sum_bio_prop_year_kg_ha, 
           richness, log_bio_kg_ha))

ch_2015_sb_veg_cs<-
  ch_2015_sb_veg%>% 
  mutate(nonative_forest_cs = scale(non.native_forest_percent, center = TRUE, scale = TRUE),
         island_size_cs = scale(island_size_ha, center = TRUE, scale = TRUE),
           rich_cs = scale(richness, center = TRUE, scale = TRUE))
str(ch_2015_sb_veg_cs)

#also combine with reduced data: 
ch_2015_bio_div_veg_R3<-inner_join(select(ch_2015_sb_veg_cs, 
                                          c("Atoll", "Island", "rat_status", "non.native_forest_percent",
                                                             "nonative_forest_cs", 
                                                            "island_size_cs")), 
                                select(ch_2015_div_R3, "Atoll", "Island", "island_size_ha", "sum_bio_prop_year_kg_ha", "log_bio_kg_ha",   "logbio_cs", "logbio_c", "richness", "rich_cs", "rich_c", ), by=c("Atoll","Island")) 

ch_2015_bio_div_veg_R3


```


#save veg data
```{r}
save(ch_2015_sb_veg, ch_2015_sb_veg_cs, 
    ch_2015_bio_div_veg_R3,
     file = "ch_2015_sb_veg_2021_09_30.Rdata")
```



###Check corrs between full data and reduced data:
```{r}
cor(ch_2015_sb_veg_cs$richness, ch_2015_bio_div_veg_R3$richness) #0.9603034
cor(ch_2015_sb_veg_cs$log_bio_kg_ha, ch_2015_bio_div_veg_R3$log_bio_kg_ha) #0.9999565
cor(ch_2015_sb_veg_cs$sum_bio_prop_year_kg_ha, ch_2015_bio_div_veg_R3$sum_bio_prop_year_kg_ha) #1
###SUPER STRONG CORRS - so excluding rare species shouldn't affect results
```


