---
title: "seabird_FD_analysis"
author: "CEB"
date: "10/20/2021"
output: html_document
---


#load package
```{r}
library(mFD)
#installed mFD package from github, but now on CRAN so updated to latest version
#https://cmlmagneville.github.io/mFD/articles/mFD_general_workflow.html
#https://github.com/CmlMagneville/mFD

library(cowplot) #for combining plots
library(tidyverse)

```



#load data
```{r}
load("ch_2015_sb_veg_2021_09_30.Rdata")

sb_trait_full_mat<-read.csv("seabird_traits_final_r.csv", stringsAsFactors = TRUE)
str(sb_trait_full_mat)

sb_trait_cats<-read.csv("seabird_traits_cat_for_mFD_r.csv", stringsAsFactors = TRUE)
str(sb_trait_cats)

#seabird species per island:
load("seabird_comm_data_2021_09_30.Rdata")
load("seabird_comm_data_REDUCED_2021_09_30.Rdata")


##seabird/site data wide:
#ch_15_bio_wide #biomass (prop year kg/ha) and breeding pairs (#/ha)

#by island, wide with diversity data:
#ch_2015_div

```


#organize seabird data
```{r}
#doesn't matter if use biomass or breeding pairs, because I'm just calculating functional richness: 
#but i think using breeding pairs makes more sense - matches community analysis
ch_2015_div_sp<-
ch_2015_div%>% 
  select(Island, Breeding_Pairs_ha_Ardenna.pacifica:Breeding_Pairs_ha_Gygis.alba ) %>% 
  rename_with( ~(gsub("Breeding_Pairs_ha_", "", .x, fixed = TRUE)))%>% 
  column_to_rownames(var = "Island")%>% 
  select(sort(colnames(.)))%>% 
  as.matrix()

ch_2015_div_sp_biomass<-
  ch_2015_div%>% 
  select(Island, Biomass_Prop_Year_kg_ha_Ardenna.pacifica:Biomass_Prop_Year_kg_ha_Gygis.alba ) %>% 
  rename_with( ~(gsub("Biomass_Prop_Year_kg_ha_", "", .x, fixed = TRUE)))%>% 
  column_to_rownames(var = "Island")%>% 
  select(sort(colnames(.)))%>% 
  as.matrix()

```



#organize and reduce trait data
```{r}
trait_mat_r<-
  sb_trait_full_mat%>% 
  select(Scientific_name, Body_mass_log_cont, 
         Breeding_duration_ord,   Site_fidelity_ordered, 
       Specific_nesting_habitat_cat, 
          Foraging_duration_ordered, 
        Foraging_habitat_ordered, 
          Fish_diet_percent, Foraging_on_or_near_surface_percent 
        ) %>%  
   mutate(Breeding_duration_ord = as.ordered(Breeding_duration_ord),
     Site_fidelity_ordered=as.ordered(Site_fidelity_ordered),
         Foraging_habitat_ordered=as.ordered(Foraging_habitat_ordered),
         Foraging_duration_ordered = as.ordered(Foraging_duration_ordered))%>%  
  arrange(Scientific_name) %>% 
    column_to_rownames(var = "Scientific_name")
  
str(trait_mat_r)

sb_trait_cats_r<-
  sb_trait_cats%>% 
  filter(trait_name == "Body_mass_log_cont" | trait_name == "Breeding_duration_ord" | 
           trait_name == "Site_fidelity_ordered" | trait_name == "Specific_nesting_habitat_cat" | 
           trait_name == "Foraging_duration_ordered" | trait_name == "Foraging_habitat_ordered" | 
            trait_name == "Fish_diet_percent" | trait_name == "Foraging_on_or_near_surface_percent") %>% 
  droplevels()
str(sb_trait_cats_r)

```



#summarize traits & communities
```{r}
#summarize traits
sb_traits_summ<-mFD::sp.tr.summary(tr_cat = sb_trait_cats_r, sp_tr = trait_mat_r)

sb_traits_summ$"tr_types"      
sb_traits_summ$"mod_list" 


# Summary of the assemblages * species dataframe:
sb_species_summ <- mFD::asb.sp.summary(asb_sp_w = ch_2015_div_sp)
head(sb_species_summ$"asb_sp_occ", 3)   #species occurrences for first 3 assemblages
sp_seabirds_occ <- sb_species_summ$"asb_sp_occ" #save species occurrences

sb_species_summ$"sp_tot_w" #species total density in all assemblages
sb_species_summ$"asb_tot_w"   #total density per assemblage
sb_species_summ$"asb_sp_richn"   #species richness per assemblage


#repeat just to look at biomass:
sb_species_summ_bio <- mFD::asb.sp.summary(asb_sp_w = ch_2015_div_sp_biomass)

sb_species_summ_bio$"sp_tot_w" #species total biomass in all assemblages
sb_species_summ_bio$"asb_tot_w"   #total biomass per assemblage
sb_species_summ_bio$"asb_sp_richn"   #species richness per assemblage
```



#calculate distances between species based on functional traits
https://cmlmagneville.github.io/mFD/reference/funct.dist.html
```{r}
sp_dist_seabird <- mFD::funct.dist(
  sp_tr         = trait_mat_r,
  tr_cat        = sb_trait_cats_r,
  metric        = "gower",
  scale_euclid  = "scale_center", #doesn't do anything when using gower distance, so may need to scale in advance
  ordinal_var   = "classic",
  weight_type   = "equal",
  stop_if_NA    = TRUE)

sp_dist_seabird
round(sp_dist_seabird, 3)    

```



#compute multidimensional functional spaces and assess their quality
```{r}
fspaces_quality_seabirds <- mFD::quality.fspaces(
  sp_dist             = sp_dist_seabird,
  maxdim_pcoa         = 10,
  deviation_weighting = c('absolute','squarred'),
  fdist_scaling       = c(TRUE,FALSE),
  fdendro             = "average")


round(fspaces_quality_seabirds$"quality_fspaces", 3)  # Quality metrics of spaces
#lowest mad = best quality, so here we see that 4d is best (with 3d close second or tied). Same regardless of whether fdist_scaling TRUE or FALSe and whether devaition_weighting absolute or squared

fspaces_quality_seabirds$details_fspaces
#so either 3 or 4 dimensions would looks good.

```



#plot quality of functional spaces
```{r}
mFD::quality.fspaces.plot(
  fspaces_quality            = fspaces_quality_seabirds,
  quality_metric             = "mad",
  fspaces_plot               = c("tree_average", "pcoa_2d", "pcoa_3d", 
                                 "pcoa_4d", "pcoa_5d", "pcoa_6d"),
  name_file                  = NULL,
  range_dist                 = NULL,
  range_dev                  = NULL,
  range_qdev                 = NULL,
  gradient_deviation         = c(neg = "darkblue", nul = "grey80", pos = "darkred"),
  gradient_deviation_quality = c(low = "yellow", high = "red"),
  x_lab                      = "Trait-based distance")

#i think based on plots 3D functional space best - looks like deviations more distributed + and -, whereas in pcoa4 most are above the 1:1 line
str(fspaces_quality_seabirds)
fspaces_quality_seabirds$pairsp_fspaces_dist 

```


#test correlation between functional axes and traits
```{r}
sp_faxes_coord_seabirds <- fspaces_quality_seabirds$"details_fspaces"$"sp_pc_coord"


seabirds_tr_faxes <- mFD::traits.faxes.cor(
  sp_tr          = trait_mat_r, 
  sp_faxes_coord = sp_faxes_coord_seabirds[ , c("PC1", "PC2", "PC3")], 
  plot           = TRUE)

#print all relationships:
seabirds_tr_faxes

# Print traits with "significant" effect:
seabirds_tr_faxes$"tr_faxes_stat"[which(seabirds_tr_faxes$"tr_faxes_stat"$"p.value" < 0.05), ]
seabirds_tr_faxes$"tr_faxes_stat"[which(seabirds_tr_faxes$"tr_faxes_stat"$"p.value" < 0.1), ]


# Return plots:
seabirds_tr_faxes$"tr_faxes_plot"

```



#plot functional space
```{r}
sp_faxes_coord_seabirds <- fspaces_quality_seabirds$"details_fspaces"$"sp_pc_coord"
sp_faxes_coord_seabirds #returns coordinates for plotting

big_plot <- mFD::funct.space.plot(
  sp_faxes_coord  = sp_faxes_coord_seabirds[ , c("PC1", "PC2", "PC3")],
  faxes           = c("PC1", "PC2", "PC3"),
  name_file       = NULL,
  faxes_nm        = NULL,
  range_faxes     = c(NA, NA),
  color_bg        = "grey95",
  color_pool      = "darkgreen",
  fill_pool       = "white",
  shape_pool      = 21,
  size_pool       = 1,
  plot_ch         = TRUE,
  color_ch        = "black",
  fill_ch         = "white",
  alpha_ch        = 0.5,
  plot_vertices   = TRUE,
  color_vert      = "blueviolet",
  fill_vert       = "blueviolet",
  shape_vert      = 23,
  size_vert       = 1,
  plot_sp_nm      = NULL,
  nm_size         = 3,
  nm_color        = "black",
  nm_fontface     = "plain",
  check_input     = TRUE)


# Plot the graph with all pairs of axes:
big_plot$patchwork
```



##repeat PCoA and trait plot so can add everything to one plot: 
```{r}
library(ape) #for pcoa
library(vegan) #for envfit

##APE MATCHES MFD, SO USE THIS***
ape_pcoa<-ape::pcoa(sp_dist_seabird)
ape_pcoa$values
#    Eigenvalues Relative_eig Rel_corr_eig Broken_stick Cum_corr_eig Cumul_br_stick
#1   1.079918952  0.578684269   0.26556645   0.24462567    0.2655665      0.2446257
#2   0.623318503  0.334010818   0.17192456   0.16770260    0.4374910      0.4123283
#3   0.250254366  0.134101049   0.09541470   0.12924106    0.5329057      0.5415693
#4   0.121613743  0.065167816   0.06903244   0.10360003    0.6019382      0.6451694

#look at rel_corr_eig because corrects for negative eigenvalues using Lingoes or Cailliez
#eigenvalues are proportional to amount of variation explained
#first 3 axes: cumulative corrected relative eigenvalues: .53
#pcoa1: .266, 2: .172, 3: .095


#extract axes and correlate with traits:
pc1_ape<-ape_pcoa$vectors[,1]
pc2_ape<-ape_pcoa$vectors[,2]
pc3_ape<-ape_pcoa$vectors[,3]

pc12d_ape <- ape_pcoa$vectors[,1:2]
pc23d_ape <- ape_pcoa$vectors[,2:3]
pc13d_ape <- ape_pcoa$vectors[,c(1,3)]

#correlate with traits
vectors_12_ape <- envfit(pc12d_ape, trait_mat_r, perm=1000, na.rm=TRUE)
vectors_12_ape

vectors_23_ape <- envfit(pc23d_ape, trait_mat_r, perm=1000, na.rm=TRUE)
vectors_23_ape

vectors_13_ape <- envfit(pc13d_ape, trait_mat_r, perm=1000, na.rm=TRUE)
vectors_13_ape


vectors_1_ape <- envfit(pc1_ape, trait_mat_r, perm=1000, na.rm=TRUE)
vectors_1_ape

vectors_2_ape <- envfit(pc2_ape, trait_mat_r, perm=1000, na.rm=TRUE)
vectors_2_ape

vectors_3_ape <- envfit(pc3_ape, trait_mat_r, perm=1000, na.rm=TRUE)
vectors_3_ape
```


#compare coordinates and vectors from ape/envfit & mfd
```{r}
#coordinates - ape
ape_sp_coord_for_plot<-
ape_pcoa$vectors%>% 
   as.data.frame() %>% 
  rownames_to_column(var = "Species")%>% 
   mutate(Species = str_replace(Species, "\\."," "))  %>% 
  select(Species:Axis.4)%>% 
  mutate(Species = as.factor(Species))
ape_sp_coord_for_plot

#coordinates - mfd 
mfd_sp_coord_for_plot<-
sp_faxes_coord_seabirds%>% 
   as.data.frame() %>% 
  rownames_to_column(var = "Species")%>% 
   mutate(Species = str_replace(Species, "\\."," "))  %>% 
  select(Species:PC3)
mfd_sp_coord_for_plot
#coordinates identical, so that's good


#correlations
seabirds_tr_faxes$"tr_faxes_stat"[which(seabirds_tr_faxes$"tr_faxes_stat"$"p.value" < 0.1), ] #mfd
#matches ape vectors from previous chunk


##all seems to match pretty well - so I think I can use mFD for r^2 and effect sizes, and use envfit and pcoa for plots. 

```



#Plot species in functional space using ggplot - mfd & ape - Axes 1 & 2
```{r}

#AXES 1&2:----
##extract arrows and scale: (scaling matches other plots)
df_vectors<-as.data.frame(vectors_12_ape$vectors$arrows*sqrt(vectors_12_ape$vectors$r)) #*ordiArrowMul(vectors_12_ape) matches output from scores(), then can multiplied by ordiArrowMul to fit on current plot. Explained in ?envfit
df_vectors$traits<-c("body mass", "% diet fish", "% foraging at surface")
df_vectors


#extract centroids for ordinal/categorical traits:
df_centroids<-as.data.frame(vectors_12_ape$factors$centroids)
df_centroids$traits<-c("seasonal breeding", "year-round breeding", 
                       "site faithful", "island faithful", "cluster faithful", "nomadic",
                       "burrows/cavities", "ground", "ground/trees", "trees/cavities", "trees", "trees/shrubs",
                       "short forays", "one adult overnight",
                       "all over", "neritic", "neritic/mid-distance", "mid-distance", "mid-distance/pelagic", "pelagic"
                      )
df_centroids$trait_types<-c("breeding duration", "breeding duration", 
                       "site fidelity", "site fidelity", "site fidelity", "site fidelity",
                       "nesting habitat", "nesting habitat", "nesting habitat", "nesting habitat", "nesting habitat", "nesting habitat",
                       "foraging duration", "foraging duration",
                       "foraging habitat", "foraging habitat", "foraging habitat", "foraging habitat", "foraging habitat", "foraging habitat"
                      )
df_centroids


#seperate categorical and ordinal traits:
df_centroids_o<-
  df_centroids%>% 
  filter(trait_types!="nesting habitat" )

df_centroids_c<-
  df_centroids%>% 
  filter(trait_types=="nesting habitat" )

#set species colours based on contributions to pca1, biomass, richness
#blue = correlations with biomass (break after top 4)
#yellow = correlations wth richness/pca1 (because same 5)
#purple = correlations with both (only A. stolidus)
cols <- c("Anous stolidus" = "#5D3A9B", "Anous tenuirostris" = "#FFC20A", 
          "Ardenna pacifica" = "gray", 
          "Fregata ariel" = "gray", "Fregata minor" = "#FFC20A", "Gygis alba" = "#0C7BDC",
          "Onychoprion anaethetus" = "#FFC20A","Onychoprion fuscatus" = "#0C7BDC", 
          "Phaethon lepturus" = "gray", 
          "Puffinus bailloni" = "gray", "Sterna dougallii" = "gray",
          "Sterna sumatrana" ="gray" ,
          "Sula leucogaster" = "gray", "Sula sula" = "#FFC20A", "Thalasseus bergii" = "#0C7BDC")

str(ape_sp_coord_for_plot)

#abbreviate species:
ape_sp_coord_for_plot2<-
   ape_sp_coord_for_plot %>% 
    mutate(Species_ab = paste(str_sub(Species, 1,1), str_extract(Species, '\\b\\w+$')))


#use 2 panels for plots because too much overlap when plot species and traits together:
test_plot_sp<-ggplot(data = ape_sp_coord_for_plot2, aes(x = Axis.1, y = Axis.2)) + 
     geom_point(aes(color = Species))+
   geom_text(aes(hjust="inward", vjust = "outward", label = Species, color = Species, fontface="italic")) + #label = Species_ab
    geom_segment(data=df_vectors,aes(x=0,xend=Axis.1,y=0,yend=Axis.2), color = "white", alpha = 0,
      arrow = arrow(length = unit(0.5, "cm"))) + 
    scale_color_manual(values = cols)+
  xlim(c(-.67, .81))+
  ylim(c(-.56, .32))+
  coord_fixed()+ #use coord_equal or coord_fixed so that way unit spacing on x and y axes are the same*
  ylab("PCoA 2")+
  xlab("PCoA 1")+
  theme_bw()+
   # guides(size = "none", colour = "none",  fill = guide_legend(title="Confidence level")) + 
    theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(),
    #            text = element_text(size = 24),
        legend.position='none'
    )
test_plot_sp

library(viridis)

test_plot_tr<-ggplot(data = ape_sp_coord_for_plot2, aes(x = Axis.1, y = Axis.2)) + 
  geom_path(data=df_centroids_o, aes(x=Axis.1,y=Axis.2,group=trait_types, colour = trait_types), lineend="round", arrow=arrow())+
geom_text(data=df_vectors,aes(x=Axis.1,y=Axis.2,label=traits,colour = traits, hjust="inward", vjust = "inward"),size=4, alpha =.8)+
  geom_text(data=df_centroids,aes(x=Axis.1,y=Axis.2,label=traits, colour = trait_types, hjust="inward", vjust = "inward"),size=4, alpha =.8)+
  geom_point(data=df_centroids_c,aes(x=Axis.1,y=Axis.2, shape=trait_types, colour = trait_types), size = 3, alpha =.8)+
  geom_segment(data=df_vectors,aes(x=0,xend=Axis.1,y=0,yend=Axis.2, colour = traits),
      arrow = arrow(length = unit(0.5, "cm"))) + 
   #   scale_color_brewer(palette="Set2")+
  scale_color_viridis(discrete=TRUE, option="D", begin = 0, end = .95)+
  xlim(c(-.67, .81))+
  ylim(c(-.56, .32))+
  coord_fixed()+
  ylab("PCoA 2")+
  xlab("PCoA 1")+
  theme_bw()+
   # guides(size = "none", colour = "none",  fill = guide_legend(title="Confidence level")) + 
    theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(),
    #            text = element_text(size = 24),
        legend.position='none'
    )
test_plot_tr


fd_plots_c1<-plot_grid(test_plot_sp, corr_plot_test,   #from community analysis file
                       test_plot_tr,
           ncol=2, rel_widths =c(2,1))

fd_plots_c1

#fd_plots<-plot_grid(test_plot_sp,corr_plot_test, test_plot_tr, #corr_plot_test from community analysis file as placeholder
       #    ncol=2, rel_widths =c(2,1))
#fd_plots


#ggsave(filename = "fd_plots_jan2022.pdf", 
#       plot = fd_plots,
#       width = 8,
#       height = 12,
#       units = "in",
#       dpi = 300,
#       family = "sans",
#       useDingbats=FALSE)

```


#Plot additional axes for supplement - 2&3
```{r}

#abbreviate species:
ape_sp_coord_for_plot2<-
   ape_sp_coord_for_plot %>% 
    mutate(Species_ab = paste(str_sub(Species, 1,1), str_extract(Species, '\\b\\w+$')))

##axis 2 & 3:----
#extract vectors:
df_vectors_23<-as.data.frame(vectors_23_ape$vectors$arrows*sqrt(vectors_23_ape$vectors$r)*.5) #can scale down to match plot area (based on envfit help file)
df_vectors_23$traits<-c("body mass", "% diet fish", "% foraging at surface")
df_vectors_23

#only keep sig traits
df_vectors_23_r<-
  df_vectors_23%>% 
  filter(traits!="% diet fish")


#extract centroids for ordinal/categorical traits:
df_centroids_23<-as.data.frame(vectors_23_ape$factors$centroids)
df_centroids_23$traits<-c("seasonal breeding", "year-round breeding", 
                       "site faithful", "island faithful", "cluster faithful", "nomadic",
                       "burrows/cavities", "ground", "ground/trees", "trees/cavities", "trees", "trees/shrubs",
                       "short forays", "one adult overnight",
                       "all over", "neritic", "neritic/mid-distance", "mid-distance", "mid-distance/pelagic", "pelagic"
                      )
df_centroids_23$trait_types<-c("breeding duration", "breeding duration", 
                       "site fidelity", "site fidelity", "site fidelity", "site fidelity",
                       "nesting habitat", "nesting habitat", "nesting habitat", "nesting habitat", "nesting habitat", "nesting habitat",
                       "foraging duration", "foraging duration",
                       "foraging habitat", "foraging habitat", "foraging habitat", "foraging habitat", "foraging habitat", "foraging habitat"
                      )
df_centroids_23

#remove non-sig traits:
df_centroids_23r<-
  df_centroids_23%>% 
  filter(trait_types=="nesting habitat"|trait_types=="breeding duration")
df_centroids_23r

#seperate ordinal and categorical:
df_centroids_23r_o<-
  df_centroids_23r%>% 
  filter(trait_types=="breeding duration")

df_centroids_23r_c<-
  df_centroids_23r%>% 
  filter(trait_types=="nesting habitat")


#set species colours based on contributions to pca1, biomass, richness
#blue = correlations with biomass (break after top 4)
#yellow = correlations wth richness/pca1 (because same 5)
#purple = correlations with both (only A. stolidus)
cols <- c("Anous stolidus" = "#5D3A9B", "Anous tenuirostris" = "#FFC20A", 
          "Ardenna pacifica" = "gray", 
          "Fregata ariel" = "gray", "Fregata minor" = "#FFC20A", "Gygis alba" = "#0C7BDC",
          "Onychoprion anaethetus" = "#FFC20A","Onychoprion fuscatus" = "#0C7BDC", 
          "Phaethon lepturus" = "gray", 
          "Puffinus bailloni" = "gray", "Sterna dougallii" = "gray",
          "Sterna sumatrana" ="gray" ,
          "Sula leucogaster" = "gray", "Sula sula" = "#FFC20A", "Thalasseus bergii" = "#0C7BDC")

str(ape_sp_coord_for_plot)


#use 2 panels for plots because too much overlap when plot species and traits together:
test_plot_sp_23<-ggplot(data = ape_sp_coord_for_plot2, aes(x = Axis.2, y = Axis.3)) + 
     geom_point(aes(color = Species))+
   geom_text(aes(hjust="inward", vjust = "outward", label = Species_ab, color = Species, fontface="italic")) + #label = Species_ab
    scale_color_manual(values = cols)+
  xlim(c(-.45, .32))+
  ylim(c(-.42, .35))+
  coord_fixed()+
  xlab("PCoA 2")+
  ylab("PCoA 3")+
  theme_bw()+
   # guides(size = "none", colour = "none",  fill = guide_legend(title="Confidence level")) + 
    theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(),
    #            text = element_text(size = 24),
        legend.position='none'
    )
test_plot_sp_23


test_plot_tr_23<-ggplot(data = ape_sp_coord_for_plot, aes(x = Axis.2, y = Axis.3)) + 
  geom_path(data=df_centroids_23r_o, aes(x=Axis.2,y=Axis.3,group=trait_types, colour = trait_types), lineend="round", arrow=arrow())+
geom_text(data=df_vectors_23_r,aes(x=Axis.2,y=Axis.3,label=traits,colour = traits, hjust="inward", vjust = "inward"),size=4, alpha =.8)+
  geom_text(data=df_centroids_23r,aes(x=Axis.2,y=Axis.3,label=traits, colour = trait_types, hjust="inward", vjust = "inward"),size=4, alpha =.8)+
  geom_point(data=df_centroids_23r_c,aes(x=Axis.2,y=Axis.3, colour = trait_types), size = 3, alpha =.8)+
  geom_segment(data=df_vectors_23_r,aes(x=0,xend=Axis.2,y=0,yend=Axis.3, colour = traits),
      arrow = arrow(length = unit(0.5, "cm"))) + 
   #   scale_color_brewer(palette="Set2")+
  scale_color_viridis(discrete=TRUE, option="D", begin = 0, end = .95)+
  xlim(c(-.45, .32))+
  ylim(c(-.42, .35))+
  coord_fixed()+
  xlab("PCoA 2")+
  ylab("PCoA 3")+
  theme_bw()+
   # guides(size = "none", colour = "none",  fill = guide_legend(title="Confidence level")) + 
    theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(),
    #            text = element_text(size = 24),
        legend.position='none'
    )
test_plot_tr_23



##combine plots and save------

supp_fd_plots<-plot_grid(test_plot_sp_23, test_plot_tr_23,
                         ncol=1, align="hv")

supp_fd_plots

#ggsave(filename = "supp_fd_plots_FigS5.pdf", 
#       plot = supp_fd_plots,
#       width = 8,
#       height = 8,
#       units = "in",
#       dpi = 300,
#       family = "sans",
#       useDingbats=FALSE)



```



##compute functional diversity indices and plot them
#alpha diversity in multidimensional space
```{r}
alpha_fd_indices_fruits <- mFD::alpha.fd.multidim(
  sp_faxes_coord   = sp_faxes_coord_seabirds[ , c("PC1", "PC2", "PC3")],
  asb_sp_w         = ch_2015_div_sp,
  ind_vect         = c("fdis", "fmpd", "fnnd", "feve", "fric", "fdiv", "fori", 
                       "fspe", "fide"),
  scaling          = TRUE,
  check_input      = TRUE,
  details_returned = TRUE)

#need to remove Ile Anglaise_PB because not enough species (# species must be < functional axes)
ch_2015_div_sp_red<-
  ch_2015_div_sp%>% 
  as.data.frame() %>% 
  rownames_to_column("Island")%>% 
  filter(Island != "Ile Anglaise_PB")%>% 
  column_to_rownames(var = "Island")%>% 
  as.matrix()


alpha_fd_indices_seabird <- mFD::alpha.fd.multidim(
  sp_faxes_coord   = sp_faxes_coord_seabirds[ , c("PC1", "PC2", "PC3")],
  asb_sp_w         = ch_2015_div_sp_red,
  ind_vect         = c("fdis", "fmpd", "fnnd", "feve", "fric", "fdiv", "fori", 
                       "fspe", "fide"),
  scaling          = TRUE,
  check_input      = TRUE,
  details_returned = TRUE)


#save some useful info: 
fd_ind_values_seabird<- alpha_fd_indices_seabird$"functional_diversity_indices"
fd_ind_values_seabird

#and details for plotting:
details_list_seabird <- alpha_fd_indices_seabird$"details"



plots_alpha <- mFD::alpha.multidim.plot(
  output_alpha_fd_multidim = alpha_fd_indices_seabird,
  plot_asb_nm              = c("Eagle Island", "Nelson's Island"), #can only plot up to 2 assemblages at a time
  ind_nm                   = c("fdis", "fide", "fnnd", "feve", "fric", 
                               "fdiv", "fori", "fspe"),
  faxes                    = NULL,
  faxes_nm                 = NULL,
  range_faxes              = c(NA, NA),
  color_bg                 = "grey95",
  shape_sp                 = c(pool = 3, asb1 = 21, asb2 = 21),
  size_sp                  = c(pool = 0.7, asb1 = 1, asb2 = 1),
  color_sp                 = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  color_vert               = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_sp                  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_vert                = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  color_ch                 = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_ch                  = c(pool = "white", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  alpha_ch                 = c(pool = 1, asb1 = 0.3, asb2 = 0.3),
  shape_centroid_fdis      = c(asb1 = 22,  asb2 = 24),
  shape_centroid_fdiv      = c(asb1 = 22,  asb2 = 24),
  shape_centroid_fspe      = 23,
  color_centroid_fspe      = "black",
  size_sp_nm               = 3, 
  color_sp_nm              = "black",
  plot_sp_nm               = NULL,
  fontface_sp_nm           = "plain",
  save_file                = FALSE,
  check_input              = TRUE) 

plots_alpha$"fric"$"patchwork"

```

##check corr between fric and richness
```{r}
fd_ind_values_seabird
cor(fd_ind_values_seabird$sp_richn, fd_ind_values_seabird$fric) #0.9288864
```


#check community patterns when remove rare species
```{r}
ch_2015_div_sp_R3<-
ch_2015_div_R3%>% 
  select(Island, Breeding_Pairs_ha_Ardenna.pacifica:Breeding_Pairs_ha_Gygis.alba ) %>% 
  rename_with( ~(gsub("Breeding_Pairs_ha_", "", .x, fixed = TRUE)))%>% 
  column_to_rownames(var = "Island")%>% 
  select(sort(colnames(.)))%>% 
  as.matrix()

ch_2015_div_sp_biomass_R3<-
  ch_2015_div_R3%>% 
  select(Island, Biomass_Prop_Year_kg_ha_Ardenna.pacifica:Biomass_Prop_Year_kg_ha_Gygis.alba ) %>% 
  rename_with( ~(gsub("Biomass_Prop_Year_kg_ha_", "", .x, fixed = TRUE)))%>% 
  column_to_rownames(var = "Island")%>% 
  select(sort(colnames(.)))%>% 
  as.matrix()


# Summary of the assemblages * species dataframe:
sb_species_summ_R3 <- mFD::asb.sp.summary(asb_sp_w = ch_2015_div_sp_R3)

sb_species_summ_R3$"sp_tot_w" #species total density in all assemblages
sb_species_summ_R3$"asb_tot_w"   #total density per assemblage
sb_species_summ_R3$"asb_sp_richn"   #species richness per assemblage


#repeat just to look at biomass:
sb_species_summ_bio_R3 <- mFD::asb.sp.summary(asb_sp_w = ch_2015_div_sp_biomass_R3)

sb_species_summ_bio_R3$"sp_tot_w" #species total biomass in all assemblages
sb_species_summ_bio_R3$"asb_tot_w"   #total biomass per assemblage
sb_species_summ_bio_R3$"asb_sp_richn"   #species richness per assemblage

```

